---
title: "Homework 4"
author: "Connor Johnson"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r packages}
library(ISLR)
library(dplyr)
library(MASS)
library(class)
library(bootstrap)
library(boot)
library(broom)
library(coin)
library(rcompanion)
```

## Problem 2

We can use the bootstrap approach to approximate the standard deviation. The bootstrap method works by re-sampling the observations from our data a large number of times. We can then use these samples to create a distribution of estimates for our prediction. Finding the standard deviation of this distribution will be an estimate for the standard deviation of out prediction.



## Problem 3

```{r 3a}
library(bootstrap)
data(law)
plot(law)
cor(law)
```

There does appear to be a strong positive linear relationship between LSAT score and GPA. The correlation between them is 0.7764

```{r 3b}


nBoot <- 1000
data(law)
mean.boot <- rep(0,nBoot)
cor.boot = numeric(nBoot)
for (i in 1:nBoot) {
	xperm1 <- sample(1:nrow(law), size = nrow(law), replace=T)
	cor.boot[i] = cor(law[xperm1,1],law[xperm1,2])
}
hist(cor.boot, breaks=20)
set.seed(5)

cor0 = cor(law$LSAT,law$GPA)
abline(v=cor0, col="red", lwd=2)

```



```{r 3cd}
set.seed(22)
boot_corr <- function(data, resample_vector) {
    cor(data$x[resample_vector], data$y[resample_vector])
}
ds <- data.frame(y = law$LSAT, x = law$GPA)
boots <- boot(ds, boot_corr, R = 1000)
boot.ci(boots)

hist(cor.boot, breaks=20)
print("Bias = -0.004255571")
abline(v=cor0, col="red", lwd=2)
abline(v=0.4490, col="blue", lwd=2)
abline(v=0.9635, col="blue", lwd=2)
abline(v=0.2958, col="green", lwd=2)
abline(v=0.9383, col="green", lwd=2)
```

(c)

Based on the confidence interval, we can't reject the null hypothesis that the correlation is equal to 0.5 because 0.5 is within the confidence interval

(d)

Based on the bias corrected confidence interval, we cannot reject the null hypothesis that correlation is equal to 0.5 because 0.t is within the confidence interval.

(e)

```{r 3e}
nperm <- 1000
law.perm <- numeric(nperm)
for (i in 1:nperm) {
  ind <- sample(nrow(law), replace = TRUE)
  law.perm[i] <- cor(law[ind, "LSAT"], law[ind, "GPA"])
}
mean(law.perm)
hist(law.perm, freq = FALSE,breaks = 20)
abline(v=cor(law$LSAT,law$GPA),col='red',lwd=2)
law.perm = sort(law.perm)
print(paste0("95% Confidence Interval: (",law.perm[25],", ",law.perm[975],")"))
```

The confidence interval does not contain zero so we reject the null hypothesis that correlation is equal to zero.



## Problem 4

(a)
```{r 4a}
X1 = runif(50)
X2 = runif(50)
epsilon = rnorm(50, 0, 0.25)
Y = X1 + X2 + epsilon
```

(b)
```{r 4b}
#set.seed(1)
X1 = runif(50)
X2 = runif(50)
epsilon = rnorm(50, 0, 0.25)
Y = X1 + X2 + epsilon
reg = lm(Y~X1+X2)
X1_test = runif(30)
X2_test = runif(30)
epsilon_test = rnorm(30, 0, 0.25)
Y_test = X1_test + X2_test + epsilon_test

summary(reg)
test = data.frame(X1_test,X2_test)
colnames(test) <- c("X1", "X2")
pred = predict(reg, test)
MSE0 = mean((pred-Y_test)^2)
```



(c)
```{r 4c}
perms = 1000
MSE = numeric(perms)
for(i in 1:perms){
  test$X1 = sample(test$X1)
  test$X2 = sample(test$X2)
  pred = predict(reg, test)
  MSE[i] = mean((pred-Y_test)^2)
  
}
hist(MSE, xlim=c(0,0.6))
abline(v=MSE0, col="blue")
print(length(MSE[MSE < MSE0])/perms)
```

We can reject the null hypothesis that $\beta_{1} = 0$ and $\beta_{2} = 0$ because the p-value is less than 0.05


(d)
```{r 4d}
perms = 1000
MSE = numeric(perms)
for(i in 1:perms){
  test$X2 = sample(test$X2)
  pred = predict(reg, test)
  MSE[i] = mean((pred-Y_test)^2)
  
}
hist(MSE, xlim=c(0,0.5))
abline(v=MSE0, col="blue")
print(length(MSE[MSE < MSE0])/perms)
```

We can reject the null hypothesis that $\beta_{1} = 0$ because the p-value is less than 0.05

(e)
```{r 4e}
m <- matrix(runif(5000), ncol=10, nrow = 500)
Xs = data.frame(m)
colnames(Xs) <- c("X1","X2","X3","X4","X5","X6","X7","X8","X9","X10")
epsilon_2 = rnorm(500, 0, 0.25)
Y = Xs$X1+Xs$X2+Xs$X3+Xs$X4+Xs$X5+Xs$X6+Xs$X7+Xs$X8+Xs$X9+Xs$X10+epsilon_2

m <- matrix(runif(500), ncol=10, nrow = 50)
Xs_test = data.frame(m)
colnames(Xs) <- c("X1","X2","X3","X4","X5","X6","X7","X8","X9","X10")
epsilon_2_test = rnorm(50, 0, 0.25)

Y_test = Xs_test$X1+Xs_test$X2+Xs_test$X3+Xs_test$X4+Xs_test$X5+Xs_test$X6+Xs_test$X7+Xs_test$X8+Xs_test$X9+
  Xs_test$X10+epsilon_2_test

reg = lm(Y~X1+X2+X3+X4+X5+X6+X7+X8+X9+X10,data=Xs)

pred = predict(reg, Xs_test)
MSE0 = mean((pred-Y_test)^2)
```


```{r 4f}
X = c(Xs_test$X8, Xs_test$X9, Xs_test$X10)
perms = 1000
MSE = numeric(perms)
for(i in 1:perms){
  X <- sample(X)
  Xs_test$X8 = sample(Xs_test$X8)
  Xs_test$X9 = sample(Xs_test$X9)
  Xs_test$X10 = sample(Xs_test$X10)
  pred = predict(reg, Xs_test)
  MSE[i] = mean((pred-Y_test)^2)
  
}
hist(MSE, xlim=c(0,0.8))
abline(v=MSE0, col="blue")
print(length(MSE[MSE < MSE0])/perms)
```

We can reject the null hypothesis that $\beta_{8} = 0$,  $\beta_{9} = 0$, and $\beta_{10} = 0$ because the p-value is less than 0.05

## Problem 5

(a)

We would perform a z-test on the two proportions of the group with hypotheses below

$$H_{0}: p_{1}=p_{2}$$ $$H_{a}: p_{1} \neq p_{2}$$ 
$p_{1}$ is the placebo group and $p_{2}$ is the vaccinated group.
The z-stat formula is
$$Z=\frac{p_{1}-p_{2}}{\sqrt{\frac{p q}{n_{1}} + \frac{p q}{n_{2}}}}$$
Plugging in our values, we get $Z=11.8347$ with $p \approx 0$. So, we can reject the null hypothesis which shows that the vaccine is effective.

(b)

We would expect that the number of positive cases in each group would be similar to each other.

(c)

We can use the difference in proportion between the two groups as our test statistic.


```{r 5d}

vaccine = numeric(21500)
placebo = numeric(21500)
for (i in 1:8){
  vaccine[i] = 1
}
for (i in 1:162){
  placebo[i] = 1
}
t_star = mean(placebo)-mean(vaccine)
perms = 1000
t = numeric(perms)
for(i in 1:1000){
  X = numeric(43000)
  idx = sample(1:length(X), 170, replace=FALSE)
  X[idx] = 1
  vaccine = X[1:21500]
  placebo = X[21501:43000]
  t[i] = mean(placebo) - mean(vaccine)
}
hist(t, xlim=c(-0.002, 0.008))
abline(v=t_star, col="blue")
print(length(t[t_star<t])/perms)
```

None of the test values are less than the original statistic so we know that the p-value is <0.001.





